Developing a preprint culture in biology

### 2012-10-18 ms.tex
finish editing

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits associated with the practice. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precedence, letting
research lay claim to an idea thus preventing it from being ``stolen'' \cite{gin11}

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice.  This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. A preprint is simply a document
that allows ideas to spread and be discussed, it is not yet formally validated
by the peer-review system. This is why the majority of publishers do not see
arXiv and similar services as a violation of the Ingelfinger rule. Almost all of
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
The Ecological Society of America and the Genetics Society of America recently
changed their policies to allow public preprints. \emph{Nature}
responded to the rumour that they refused manuscripts submitted to arXiv by
saying that ``\emph{Nature} never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.  Still, a few
journals adopt a ``by default'' hostile attitude towards preprints, mostly due
to the lack of clear policy of the publishers, or perhaps because a preprint
culture has not developed in biology and the practice is still considered
unusual. As an example, Wiley-Blackwell, which publishes some of the leading
journals in biology, has no official policy on the matter.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Genetics Society of America                 & Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in biology. Some publishers
tolerate preprints except for a few of their medical journals, e.g.: Journal
of the National Cancer Institute from Oxford and The Lancet from Elsevier.}
    \label{table:policies}
\end{table*}

@@

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, It might significantly reduce its negative
consequences. Preprint servers speed up publication and make no judgement on
pertinence and originality. Paradoxically, because peer-reviewing is conducted
\emph{after} the manuscript is made public, preprint servers might bring it
closer to its fundamental objective. The role of peer-reviewing is to judge the
scientific quality of a study. It is the first barrier against the fraudulent
and poor quality science susceptible to impede scientific progress. Preprints
are not filtered, neither for their quality nor their pertinence.  Technically,
the difference between a preprint and a traditional publication should be that
the latter as the approval stamp by pairs on its quality. The relevance of a
study, its contribution to science, should only be judged post-publication by
many more readers than the typical two-four anonymous reviewers. With a such a
shift in the diffusion strategy, the role of traditional journals and their
editors would be to showcase scientific discoveries for specialized readership.
Such a process should improve even further their diffusion, not impede it.

>>>>

There are a number of reasons why biologists did not develop the reflex of
sharing preprints, many of which are based on common misconceptions. For
example, in contrast to other fields there is a perception in biology that
public preprints make it easier to steal ideas \cite{gin11}. However, there is
no evidence of this happening in the numerous other fields that have adopted
preprint servers, and since preprint servers create a clear record of who had
the idea first, and when, this appears to be a largely unfounded concern. In
other fields preprints serve the opposite role, they allow straightforward
establishment of precedence, letting research lay claim to an idea thus
preventing it from being ``stolen'' \cite{gin11}

Another major concern is based on a certain interpretation of the Ingelfinger
rule: scientists should not publish the same manuscript twice \cite{alt96}. A
preprint is simply a document that allows ideas to spread and be discussed, it
is not yet formally validated by the peer-review system. This is why the
majority of publishers do not see arXiv and similar services as a violation of
the Ingelfinger rule. Almost all of the major publishers in biology are
preprint-friendly, including: Nature Publishing Group, PLOS, BMC, PNAS,
Elsevier, and Springer \ref{table:policies}. This year, both the Ecological
Society of America and the Genetics Society of America changed their policies to
allow public preprints. \emph{Nature} even felt compelled to respond to the
rumour that they refused manuscripts submitted to arXiv by saying that
``\emph{Nature} never wishes to stand in the way of communication between
researchers. We seek rather to add value for authors and the community at large
in our peer review, selection and editing'' \cite{nat05}.  Still, a few journals
adopt a ``by default'' hostile attitude towards preprints, mostly due to the
lack of clear policy of the publishers, or perhaps because a preprint culture
has not developed in biology and the practice is still considered unusual. As an
example, Wiley-Blackwell, which publishes some of the leading journals in
biology, has no official policy on the matter.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Genetics Society of America                 & Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in biology. Some publishers
tolerate preprints except for a few of their medical journals, for example the
\emph{ Journal of the National Cancer Institute} from Oxford and \emph{The
Lancet} from Elsevier.}
    \label{table:policies}
\end{table*}

@@

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, It might significantly reduce its negative
consequences. The role of peer-reviewing is to judge the scientific quality of a
study. It is the first barrier against the fraudulent and poor quality science
susceptible to impede scientific progress. In practice, the peer-review system
is not only used to evaluate scientific quality but also to judge pertinence. On
the other hand, preprints are not filtered, neither for their quality nor their
pertinence. Widespread adoption of preprint servers has the potential to shift
the diffusion strategy: journals would remain important to validate
publications, but the relevance of a study, its contribution to science, should
only be judged by many more readers than the typical two-four anonymous
reviewers. With a shift in the diffusion strategy, the role of traditional
journals and their editors would be to showcase scientific discoveries for
specialized readership.
